Modelling the Retinotectal projection - can competition provide a stopping mechanism?

### 2021-11-17 paper/paper.tex
Adds letters to Fig 3

\subsection*{Simulation}

@@

@@

Fig.\,\ref{f:ch}A indicates the colour code used for retinal cell position;
dorsal retinal cells are indicated by green, nasal cells by red. Lightness
gives position along the relevant axis so that cells at the dorsal-nasal
corner are yellow and those at the ventral-transverse corner are
black. Fig.\,\ref{f:ch}E shows the expected layout of retinal ganglion cell
axons on the tectum. The axons retain the colour of their originating cell on
the retina, as shown in Fig.\,\ref{f:ch}A. Nasal cell axons should be arranged
caudally; dorsal axons laterally; axons should be evenly spaced and the
pattern should span the entire tectum.  Fig.\,\ref{f:ch}B--D show `fish net'
plots of the centroid of the branches of each axon at three timepoints. In
these plots, lines are drawn between axons that are expected to arrange
themselves adjacently to one another. The number of line crossings in this
plot gives an indication of the disorder in the arrangement. Fig.\,\ref{f:ch}G
shows the positions of the individual branches (i.e.~growth cones) at $t=300$
that contribute to the the plot in Fig.\,\ref{f:ch}D. By this time, the
topological order matches that shown in Fig.\,\ref{f:ch}E. Opposing gradients
have sorted the axons along the two axes to form a grid of cells. However, the
arrangement does not span the full tectum. Fig.\,\ref{f:ch}F shows two pattern
metrics; the sum of squared distance errors of the axon centroids and the
number of crossings in the fish net plot of Fig.\,\ref{f:ch}D. The expected
number of crossings is 0, as in Fig.\,\ref{f:ch}E. Fig.\,\ref{f:ch}H shows
selected axons (one from the centre and one from each corner of the retina)
and their position history.

@@

The result at $t=300$ is shown in Fig.\,\ref{f:chalt}A and B. The axon
centroids are arranged across the full tectal surface.  The arrangement is
slightly disordered around the middle of the tectum. This is caused by
numerical artefacts from the discretization of the ligand expression and the
fact that in the centre, the difference between the opposing receptor-ligand
signals is very small.  Fig .\,\ref{f:chalt}C shows the error metric, which is
closer to 0 with the new ligand expression.

@@

>>>>

\subsection*{Assumptions}

@@

\subsection*{Chemoaffinity alone}

@@

Results are shown in Fig.\,\ref{f:ch}. Fig.\,\ref{f:ch}A indicates the colour
code used for retinal cell position; dorsal retinal cells are indicated by
green, nasal cells by red. Lightness gives position along the relevant axis so
that cells at the dorsal-nasal corner are yellow and those at the
ventral-transverse corner are black. Fig.\,\ref{f:ch}E shows the expected
layout of retinal ganglion cell axons on the tectum. The axons retain the
colour of their originating cell on the retina, as shown in
Fig.\,\ref{f:ch}A. Nasal cell axons should be arranged caudally; dorsal axons
laterally; axons should be evenly spaced and the pattern should span the
entire tectum.  Fig.\,\ref{f:ch}B--D show `fish net' plots of the axon
centroids at three timepoints. An axon's centroid is the average location of
its 4 branches. In these plots, lines are drawn between axons that are
expected to arrange themselves adjacently to one another by the end of the
simulation. The number of line crossings in this plot gives an indication of
the disorder in the arrangement. Fig.\,\ref{f:ch}G shows the positions of the
individual branches (i.e.~growth cones) at $t=300$ that contribute to the the
plot in Fig.\,\ref{f:ch}D. By this time, the topological order matches that
shown in Fig.\,\ref{f:ch}E. Opposing gradients have sorted the axons along the
two axes to form a grid of cells. However, the arrangement does not span the
full tectum. Fig.\,\ref{f:ch}F shows two pattern metrics; the sum of squared
distance errors of the axon centroids and the number of crossings in the fish
net plot of Fig.\,\ref{f:ch}D. The expected number of crossings is 0, as in
Fig.\,\ref{f:ch}E. Fig.\,\ref{f:ch}H shows selected axons (one from the centre
and one from each corner of the retina) and their position history.

@@

for the tectal ligand expression.

The result at $t=300$ is shown in Fig.\,\ref{f:chalt}A and B. The axon
centroids are arranged across the full tectal surface.  The arrangement is
slightly disordered around the middle of the tectum. This is caused by
numerical artefacts from the discretization of the ligand expression and the
fact that in the centre, the difference between the opposing receptor-ligand
signals is very small. Nevertheless, it is demonstrated that a modified form
for the tectal ligand expression (which was found easily by hand tuning)
allows a well ordered map to form for the wildtype system. Fig
.\,\ref{f:chalt}C shows the error metric, which is closer to 0 with the new
ligand expression.

@@

The `chemoaffinity only' model begins to encounter problems when we introduce
simulations of experimental manipulations.
